### 2021-11-21 draft.tex
Deploying to gh-pages from @ gottacatchenall/thesis_proposal@cc438154ba9b06f029a0d427c6faaa9eb4a4dc47 🚀

But forecasting is only half the story. Marx's most well known quote
that ``philosophers have hitherto only interpreted the world in various
ways; the point is to change it.'' Indeed, once we have a forecast about
how an ecosystem will change in the future, what if this forecast
predicts a critical ecosystem service will deteriorate? We are still
left with the question, what do we in the time being to mitigate the
negative consequences a forecast predicts? In this framing, mitigating
the consequences of anthropogenic change on ecosystems becomes an
optimization problem: given a forecast of the probability. We have some
goal state for the future, and some estimate of what the state of the
world will be given a set of actions. Frame optimization problem
mathematically an introduce concept of solution-space and constraint.

@@

Plants and pollinators form interaction networks, called the
``architecture of biodiversity'' (\textbf{Jordano2007?}).

The stability, function, and persistance of ecosystems relies on the
structure of these interactions. Antropogenic change threatens to
unravel these networks. Two aspects to this change: spatial and
temporal. Spatially, range shifts along elevational gradient, and
temporall, phenological shifts.

The issue is that we don't really know what interactions are like now.
So not only do we need to predict with data that is spatially and
temporally sparse and likely to contain many interaction
``false-negatives'' (Strydom \emph{et al.} 2021).

This chapter uses several years of data on bee-flower phenology and
interactions, combined with spatial records of species occurrence via
GBIF, to forecast how much overlap there will be between
plants/pollinators in space/time.

@@

Split the process into parts.

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Building an interaction prediction model. 2) Make it spatial based on
  distributions. 3) Forecast distributions based on CMIP6.
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  we got a tree
\end{enumerate}

@@

There are false-negatives in interation data. Co-occurrence is not the
same thing as interaction (\textbf{cite?}), but often is used as a
proxy.

This chapter unravels the relationship between a given species relative
abundance and the sampling effort needed to adequately understand this
species distribution and interactions.

There is more than one way to observe a false-negative.

\begin{figure}
\centering
\includegraphics{./figures/ch2.png}
\caption{taxonomy of false negatives}
\end{figure}

It begins with a conceptual framework for understanding the difference
in false-negatives in occurrence, co-occurrence, and interactions
(fig.~3). We use a null model of the relative-abundance distribution
(Hubbell 2001) to simulate realized false-negatives as a function of
varying sampling effort.

This also goes on to includes testing some assumptions of the model with
empirical data fig.~\ref{fig:posassoc}. which indicate our neutral
model, if anything, underestimates the probability of false-negatives
due to positive correlations in co-occurrence in two spatially
replicated networks (Thompson \& Townsend 2000; Hadfield \emph{et al.}
2014)---further I'm planning to add the field data from chapter one into
this anlysis once complete.

@@

new addition: - simulate species distribution and efficacy of detection
given a set of observation points where the dist from observation site
decays. optimize set of repeated sampling locations L for a \emph{known}
distribution D. address SDM not being the territory

\hypertarget{results}{%
\subsection{Results}\label{results}}

\begin{itemize}
\tightlist
\item
  nonrandom association figure sampling effort under neutral model
\end{itemize}

@@

This chapter proposes a general algorithm for optimizing corridor
placement based on a measurement of ecosystem functioning derived from
simulations run on a proposed landscape modification. We propose various
landscape modifications which alter the cover of a landscape,
represented as a raster (fig.~6, left). We then compute a new resistance
surface based on the proposed landscape modification, and based on the
values of resistance to dispersal between each location we simulate
spatially-explicit metapopulation dynamics model (Hanski \& Ovaskainen
2000; Ovaskainen \emph{et al.} 2002) to estimate a distribution of time
until extinction for each landscape modification (fig.~6, right).

@@

\begin{itemize}
\tightlist
\item
  land cover -\textgreater{} resistance -\textgreater{} extinction time
  simulated annealing to
\item
  optimize landscape optimization
\end{itemize}

\hypertarget{ch4-a-software-note-on-the-resulting-packages.}{%
\section{CH4 a software note on the resulting
packages.}\label{ch4-a-software-note-on-the-resulting-packages.}}

(MetacommunityDynamics.jl: a virtual laboratory for community ecology):
a collection of modules in the Julia language for different aspects of
metacommunity ecology, including most of the code used for the preceding
chapters.

\begin{figure}
\centering
\includegraphics{./figures/ch4.png}
\caption{todo}
\end{figure}

@@

>>>>

But forecasting is only half the story---if indeed ``{[}ecologists{]}
have hitherto only interpreted the world in various ways; the point is
to change it,'' then once we have a forecast about how an ecosystem will
change in the future, what if this forecast predicts a critical
ecosystem service will deteriorate? We are still left with the question,
what do we in the time being to mitigate the negative consequences a
forecast predicts? In this framing, mitigating the consequences of
anthropogenic change on ecosystems becomes an optimization problem:
given a forecast of the probability. We have some goal state for the
future, and some estimate of what the state of the world will be given a
set of actions. Frame optimization problem mathematically an introduce
concept of solution-space and constraint.

@@

Interactions between plants and pollinators form networks of
interactions, which structure the ``architecture of biodiversity''
(\textbf{Jordano2007?}). The persistence, functioning, and stability of
ecosystems emerge from these interactions, but antropogenic change
threatens to unravel and ``rewire'' these networks
(\textbf{CaraDonna2017IntRew?}). Plant-pollinator networks face two
possible forms of rewiring in response to anthropogenic environmental
change: spatial and temporal.\\
Spatially, range shifts could cause interacting species to no longer
overlap in space, and shifts in phenology could cause interacting
species to no longer overlap in time.

This chapter uses several years of data on bumblebee-flower phenology
and interactions across several field site, each consisting of several
plots across an elevational gradient, combined with spatial records of
species occurrence via GBIF to forecast this uncoupling.

Addresses the EBV to forecast of EBV element of the flow from data to
mitigation in fig.~\ref{fig:thesis}.

@@

The issue is that we don't really know what interactions are like now.
So we also need to predict interactions as they exist \emph{in the
present} as the data we have and temporally sparse and likely to contain
many interaction ``false-negatives'' (Strydom \emph{et al.} 2021).

The process of going from data to forecast can be split into the
following parts

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Building an interaction prediction model (or rather a set of candidate
  models, relative abundance, phylo embedding relative abundance + phylo
  embedding) a la Strydom \emph{et al.} (2021)
\end{enumerate}

Reconstructing latent features for each species based on simulating
trait evolution on a phylogeny (\textbf{Strydom2021FooWeb?}).

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\item
  Make it spatial based on distributions.
\item
  Forecast distributions based on CMIP6
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  we got a tree and SDMs
\end{enumerate}

@@

This chapter uses simulation models to investigate the relationship
between species relative abundance, sampling effort, and probability of
accurately detecting an interaction between species, and further
proposes a method for optimizing the spatial sampling locations to
maximize the probability of detecting an interaction between two species
given their distributions. This addresses the optimization of monitoring
network part of the flow from data to mitigation in
fig.~\ref{fig:thesis}.

As explored in the previous chapter, there are false-negatives in
interation data. There is more than one way to observe a false-negative
when sampling interactions: (fig.~\ref{fig:fnrtaxonomy}). It begins with
a conceptual framework for understanding the difference in
false-negatives in occurrence, co-occurrence, and interactions.
Co-occurrence is not the same thing as interaction (Blanchet \emph{et
al.} 2020), but often is used as a proxy.

\begin{figure}
\hypertarget{fig:fnrtaxonomy}{%
\centering
\includegraphics{./figures/ch2.png}
\caption{taxonomy of false negatives}\label{fig:fnrtaxonomy}
}
\end{figure}

We use a log-normal distribution as a null model of the
relative-abundance distribution (Hubbell 2001) to simulate realized
false-negative rate as a function of varying sampling effort.

This also goes on to includes testing some assumptions of the model with
empirical data fig.~\ref{fig:posassoc}, which we analytically show that
our neutral model, if anything, underestimates the probability of
false-negatives due to positive correlations in co-occurrence in two
spatially replicated networks (Thompson \& Townsend 2000; Hadfield
\emph{et al.} 2014)---further I'm planning to add the field data from
the previous chapter into this anlysis once available.

@@

Finally this chapter proposes a simulated annealing method to optimize
the efficacy of interactoin detection given a set of observation points
where the dist from observation site decays. optimize set of repeated
sampling locations L for a pair of species \emph{known} distributions
\(D_a, D_b\).

@@

This chapter proposes a general algorithm for choosing corridor
placement to optimize a measurement of ecosystem functioning derived
from simulations run on each proposed landscape modification.

We propose various landscape modifications which alter the cover of a
landscape, represented as a raster (fig.~6, left). We then compute a new
resistance surface based on the proposed landscape modification, and
based on the values of resistance to dispersal between each location we
simulate spatially-explicit metapopulation dynamics model (Hanski \&
Ovaskainen 2000; Ovaskainen \emph{et al.} 2002) to estimate a
distribution of time until extinction for each landscape modification
(fig.~6, right).

@@

\begin{itemize}
\item
  land cover -\textgreater{} resistance -\textgreater{} extinction time
\item
  brief overview of simulated annealing
\item
  describe how you build the proposal function
\item
  optimize landscape optimization
\end{itemize}

\hypertarget{chapter-four-metacommunitydynamics.jl-a-virtual-laboratory-for-community-ecology}{%
\section{Chapter Four: MetacommunityDynamics.jl: a virtual laboratory
for community
ecology}\label{chapter-four-metacommunitydynamics.jl-a-virtual-laboratory-for-community-ecology}}

This chapter consists of a collection of modules in the Julia language
for different aspects of community ecology, including most of the code
used for the preceding chapters.

Indeed \texttt{MetacommunityDynamics.jl} is the center of this, but due
to the nature of the Julia language, it is interoperable with serveral
existing packages.

\texttt{EcoJulia} organization, including several which I have
contributed to enable interoperability with MCD.jl.

A diagram showing the relation between these packages is shown in
fig.~\ref{fig:software}.

\begin{figure}
\hypertarget{fig:software}{%
\centering
\includegraphics{./figures/ch4.png}
\caption{todo}\label{fig:software}
}
\end{figure}

@@

\leavevmode\hypertarget{ref-Blanchet2020CooNot}{}%
Blanchet, F.G., Cazelles, K. \& Gravel, D. (2020). Co-occurrence is not
evidence of ecological interactions. \emph{Ecology Letters}, 23,
1050--1063.
